---
title: "Opihi_LarvaeSurvival"
author: "Gabi Mukai"
date: "2024-01-16"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(googlesheets4)
library(colorspace)
survival <- read_sheet('https://docs.google.com/spreadsheets/d/10z96KBKw9umDFoErzKku8eiQMh4RgWS2MIzMLrLX6CY/edit#gid=361953853', "Survival Summary")
library(ggeffects)
library(ggplot2)
library(dplyr)
ggplot(survival) + 
  geom_point(aes(x = Days_in_TB, y = Proportion, color = as.factor(Treatment), shape = as.factor(Exp))) +scale_x_continuous(breaks = seq(1,12,1)) # each point is it's own female/rep

ggplot(survival) + 
  geom_point(aes(x = Days_in_TB, y = Proportion_Original, color = as.factor(Treatment), shape = as.factor(Exp))) +scale_x_continuous(breaks = seq(1,12,1)) # each point is it's own female/rep

ggplot(survival) + 
  geom_point(aes(x = Days_in_TB, y = Proportion, color = as.factor(Treatment), shape = as.factor(Exp))) +scale_x_continuous(breaks = seq(1,12,1)) + geom_smooth(aes(x = Days_in_TB, y = Proportion))

survival %>% group_by(Treatment, Days_in_TB) %>% reframe(live_total = sum(Number), original = sum(Original_Starting))


```
per temp, 5 females with at least each having 5 (for total of 25 larvae per temp)

# Get Variables Ready
```{r}
# quick check
library(stringi)
length(unique(survival$Female)) == length(unique(paste(stri_sub(survival$Female_OriginalID, length = 1),survival$Exp)))


survival$Female <- as.factor(survival$Female)
survival$Alkaline <- as.factor(survival$Alkaline)
survival$Days_in_TB <- as.numeric(survival$Days_in_TB)
survival$Treatment <- as.factor(survival$Treatment)
survival$Temperature <- 32.5
survival$Temperature[which(survival$Treatment == 9)] = 29.5
survival$Temperature[which(survival$Treatment == 6)] = 25
survival$Temperature[which(survival$Treatment == 10)] = 31
survival$Temperature[which(survival$Treatment == 1)] = 17.5
class(survival$Temperature)
class(survival$Rep)
survival$Rep <- as.factor(survival$Rep)
survival$Exp <- as.factor(survival$Exp)
survival$Original_Starting <- as.integer(survival$Original_Starting)
survival$Number <- as.integer(survival$Number)

#survival = survival[
#  with(survival, order(Treatment, Female, Alkaline, Rep, Days_in_TB)),
#]

#for (i in 2:nrow(survival)){ifelse(survival$Previous_Starting[i] != survival$Number[i-1], print(paste0("mismatch row",i)), print(("all good")))} # double check that previous value corresponds to number in previous row 
```
Let's look at our numbers 

```{r}
subset(survival, Days_in_TB == 1) %>% select(Temperature, Female, Original_Starting) %>%  group_by(Temperature) %>% summarise(number_per_treatment = sum(Original_Starting))

subset(survival, Days_in_TB == 1) %>% dplyr::select(Temperature, Female, Original_Starting, Exp) %>%  group_by(Exp, Temperature) %>% summarise(number_per_treatment = sum(Original_Starting))

OG = survival %>% group_by(Temperature, Days_in_TB) %>% summarise(Total_Previous = sum(Previous_Starting))

aggregated = survival %>% group_by(Temperature, Days_in_TB) %>% summarise(Number_Total = sum(Number)) 
#color = sequential_hcl(5,palette = "Hawaii")[5:1]
color = c("#B0F4FA","#5DD291" ,"#9A9800", "#CC3A29", "#8B0069")
raw_daily = merge(OG, aggregated, by = c("Temperature", "Days_in_TB") )%>% 
  ggplot() + 
   geom_point(aes(x = Days_in_TB, y = Number_Total/Total_Previous, color = (Temperature)), position = position_jitter(width = 0.1), size = 3)  + scale_x_continuous("Days", breaks = seq(1, 17, by = 1)) + scale_y_continuous(breaks = seq(0,1,0.25)) + theme_classic() + ylab("Daily Proportion of Survivors") + scale_color_manual(values = color[1:5]) + labs(color =  expression("Temperature " ( degree~C))) 


x = merge(OG, aggregated, by = c("Temperature", "Days_in_TB") )
x$proportion <- x$Number_Total/x$Total_Previous
x = x %>% group_by(Temperature, Days_in_TB) %>% reframe(daily_survival = proportion, cumulative_probability = NA)
x$cumulative_probability[which(x$Days_in_TB == 1)] <- x$daily_survival[which(x$Days_in_TB == 1)]
x =x %>%
  arrange(Temperature, Days_in_TB) %>%  # Ensure data is sorted by temperature and day
  group_by(Temperature) %>%  # Group by temperature treatment
  mutate(cumulative_probability = cumprod(daily_survival))
x$Temperature <- as.factor(x$Temperature)
ggplot(x) + 
   geom_point(aes(x = Days_in_TB, y = cumulative_probability, color = (Temperature)), position = position_jitter(width = 0.1), size = 3)  + scale_x_continuous("Days", breaks = seq(1, 17, by = 1)) + scale_y_continuous(breaks = seq(0,1,0.25)) + theme_classic() + ylab("Daily Proportion of Survivors") + scale_color_manual(values = color[1:5]) + labs(color =  expression("Temperature " ( degree~C))) 


OG = survival %>% group_by(Temperature, Days_in_TB) %>% summarise(Total_Original = sum(Original_Starting)) %>% subset(Days_in_TB == 1)

#aggregated = survival %>% group_by(Temperature, Days_in_TB) %>% summarise(Number_Total = sum(Number)) 
#color = sequential_hcl(5,palette = "Hawaii")[5:1]
raw_cumulative = merge(OG, aggregated, by = c("Temperature") ) %>% 
  ggplot() + 
   geom_point(aes(x = Days_in_TB.y, y = Number_Total/Total_Original, color = (Temperature)),position = position_jitter(width = 0.1), size = 3)  + scale_x_continuous("Days", breaks = seq(1, 17, by = 1)) + scale_y_continuous(breaks = seq(0,1,0.25)) + theme_classic() + ylab("Cumulative Proportion of Survivors") + scale_color_manual(values = color[1:5]) + labs(color =  expression("Temperature " ( degree~C))) 


#ggsave("~/Desktop/Opihi_Larvae/opihi_exp3raw.png")
```
# Make our Model (glmer)

```{r}
library(lme4)
# full model with response variable: Number, Original-Number
# Random effects for female, rep, alkaline, experiment trial
# fixed effects for days in thermal block and temperature
# using a binomial family
mod = (glmer( cbind(Number,Previous_Starting-Number) ~ (1 | Female/Rep)  + (1 | Alkaline) + (1 | Exp)  + Days_in_TB+Temperature, family = binomial(link="logit"), data = survival ))
summary(mod) # age matters 
AIC(mod)
library(sjPlot)
plot_model(mod, type = "pred", terms = c("Days_in_TB","Temperature"))
confint(mod)
# stay with large model 

```
# Test for Overdispersion
overdispersion function I got from https://bbolker.github.io/mixedmodels-misc/glmmFAQ.html#overdispersion
```{r}
overdisp_fun <- function(model) {
rdf <- df.residual(model)
rp <- residuals(model,type="pearson")
Pearson.chisq <- sum(rp^2)
prat <- Pearson.chisq/rdf
pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}
overdisp_fun(mod) # yes overdispersion (I think)
# dealing with overdispersion (likely won't have a huge impact but still deal with it)
install.packages("glmmTMB")
library(glmmTMB)

#repnested = (glmmTMB( cbind(Number,Previous_Starting-Number) ~ (1 | Female/Rep) + Days_in_TB + Temperature+ I(Temperature^2), family = betabinomial, data = survival )) # dropped alk as random effect bc variance were very small and messing up code
survival$Temperature = as.factor(survival$Temperature)
repnested = (glmmTMB( cbind(Number,Previous_Starting-Number) ~ (1 | Female/Rep) + Days_in_TB*Temperature, family = betabinomial, data = survival ))
cum_repnested = (glmmTMB( cbind(Number,Original_Starting-Number) ~ (1 | Female/Rep) + Days_in_TB*Temperature, family = betabinomial, data = survival ))
# use temp as factor since effect of temp is not cont. (big drop off at 32.5)
AIC(repnested)
summary(repnested)


```
# Model Diagnostics 
How do we do model diagnostics for binomial model? 

```{r}

hist(resid(repnested)) # normal 
plot(fitted(repnested),resid(repnested)) # a little skewed but okay

hist(resid(cum_repnested)) # normal 
plot(fitted(cum_repnested),resid(cum_repnested)) 
```
# Create effects plots
use ggpredict or ggeffect? 
output of ggpredict and ggeffect have the same structure I think so should be pretty easy to go between the two if needed and their output is pretty similar
```{r}
g <- ggpredict(repnested, terms=c("Days_in_TB [all]", "Temperature"))
gcum <- ggpredict(cum_repnested, terms=c("Days_in_TB [all]", "Temperature"))
#color = sequential_hcl(5,palette = "Hawaii")[5:1]

plot_mod = ggplot(g, aes(x, predicted)) +
  geom_ribbon(aes(ymin=conf.low, ymax=conf.high, fill=group, alpha=group)) + 
  scale_alpha_manual(values = c("17.5" = 0.15, "25" = 0.15, "29.5"=0.15, "31" = 0.15, "32.5" = 0.02), guide="none") +  
  geom_line(aes(color=group), linewidth = 2.5, linetype="dashed") +
  scale_color_manual(values=c(color))+
  scale_fill_manual( values=c(color), guide="none") +
  scale_x_continuous(breaks = seq(min(g$x), max(g$x), by = 1))+
  xlab(label = "Days")+
  ylab("Probability of Daily Survival") + labs(color =  expression("Temperature " ( degree~C))) + theme_classic()

ggplot(gcum, aes(x, predicted)) +
  geom_line(aes(color=group), linewidth = 1, linetype="dashed") +
  geom_ribbon(aes(ymin=conf.low, ymax=conf.high, fill=group), alpha=0.15) + 
  scale_color_manual(values=c(color))+
  scale_fill_manual( values=c(color), guide="none") +
  xlab(label = "Days")+
  ylab("Probability of Daily Survival") + labs(color =  expression("Temperature " ( degree~C))) + theme_classic()

library(patchwork)
raw_cumulative + raw_daily 
ggsave("~/Desktop/Opihi_Larvae/opihi_exp3raw.png")

plot_mod
ggsave("~/Desktop/Opihi_Larvae/opihi_exp3predicted.png")


```

# Test for significance on specific days 
what are differences in daily survival proportion on the first day, 3 day? 
```{r}
library(emmeans)

emmeans(repnested, pairwise ~ Temperature|Days_in_TB, at = list(Days_in_TB= c(1,2)), adjust = "tukey") 
# differences were tested for significance at day 1 and 3 by using library emmeans adjusted with tukey (tukey controls for doing multiple tests)
# eg difference between temp 1 and 2 on day 1 sign or not

emmeans(cum_repnested, pairwise ~ Temperature|Days_in_TB, at = list(Days_in_TB= c(1,2)), adjust = "tukey")
```
KM survival analysis? 

```{r}
# covert data frame to reflect individuals as rows over time and alive/dead as 1/0's 
binary_surival = purrr::pmap_dfr(as.data.frame(survival) %>% dplyr::select(c(Female, Temperature, Alkaline, Rep, Days_in_TB, Exp, Number, Previous_Starting)), 
              function(Female, Temperature, Alkaline, Rep, Days_in_TB, Exp, Number, Previous_Starting) {
                           data.frame(Female = Female,
                                      Temperature = Temperature,
                                      Alkaline = Alkaline,
                                      Rep = Rep,
                                      Days_in_TB = Days_in_TB,
                                      Exp = Exp,
                                      status = c( rep(0, Number),# alive
                                                rep(1, Previous_Starting - Number) ) )
                      }
              )

nrow(binary_surival) == sum(survival$Previous_Starting)
library(survival)
library(ggsurvfit)
library(coxme)
#Surv <- Surv(time = binary_surival$Days_in_TB, event = binary_surival$status)
binary_surival$Temperature <- as.factor(binary_surival$Temperature)
KM <- survfit2(Surv(time = Days_in_TB, event = status) ~  Temperature, data = binary_surival) 
cox <- coxme(Surv(time = Days_in_TB, event = status) ~  (1 | Female/Rep) + Temperature, data = binary_surival)
library(survminer)
ggsurvfit(KM) +
  labs(
    x = "Days",
    y = "Overall survival probability"
  )+ 
  add_confidence_interval()


survdiff(Surv(time = Days_in_TB, event = status) ~ Temperature, data = binary_surival)
pairwise_survdiff(Surv(time = Days_in_TB, event = status) ~ Temperature, data = binary_surival)
summary(KM, times = c(1,2))
summary(cox)


emmeans(cox, pairwise ~ Temperature, at = list(Days_in_TB= c(1)), adjust="mvt") 
emmeans(cox, pairwise ~ Temperature*Days_in_TB, at = list(Days_in_TB = c(1, 2)))

d1 <- binary_surival %>% subset(Days_in_TB%in% c(1,2)) %>% dplyr::select(Temperature, Days_in_TB, status) %>% table() %>% melt()
pairwise.prop.test(d1$value[1:10], (d1$value[1:10]+d1$value[11:20]), p.adjust.method = "bonferroni")

```
 How long did larvae live for?

```{r}
PLD_25 = survival %>% subset(Treatment == 6) %>% group_by(Female, Alkaline, Rep) %>% summarise(dead =(Previous_Starting - Number), Age = Days_in_TB)
mean(rep(PLD_25$Age,PLD_25$dead))
median(rep(PLD_25$Age,PLD_25$dead))
hist(rep(PLD_25$Age,PLD_25$dead), main = "Frequency of Larvae's Ages", xlab = "Age (Last Seen Alive)", breaks = 17, xlim = c(1,18))
ggplot() +
  geom_histogram(aes(rep(PLD_25$Age,PLD_25$dead)), binwidth = 1, color = "grey9", fill = "darkblue") + theme_bw() + xlab("Age (Day Last Seen Alive)") + ylab("Frequency")+
  scale_x_continuous(breaks = seq(min(PLD_25$Age), max(PLD_25$Age), by = 1))

p = survival %>% subset(Treatment == 6) %>% group_by(Female, Alkaline, Rep) %>% summarise(alive =(Number), Age = (Days_in_TB + 1) )

hist(rep(p$Age,p$alive), main = "Number of Live Larvae Each Day", xlab = "Age", breaks = 16, xlim = c(2,17))
ggplot() +
  geom_histogram(aes(rep(p$Age,p$alive)), binwidth = 1, color = "grey9", fill = "darkblue") + theme_bw() + xlab("Age (Days)") + ylab("Frequency")+
  scale_x_continuous(breaks = seq(min(p$Age), max(p$Age)-1, by = 1))

survival %>% subset(Treatment == 6 & Days_in_TB == 1) %>% reframe(total_start= sum(Original_Starting))
table(rep(p$Age,p$alive)) # number alive at age x 
table(rep(p$Age,p$alive))/73 # proportion alive 


```


```{r}
